P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2025-04-30 sage_latex_template/manuscript.tex
minor
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

% Changes during review
\usepackage{soul}
\usepackage{xcolor}

%\DeclareRobustCommand{\RMchange}[1]{{\sethlcolor{cyan}\hl{#1}}}
%\newcommand{\RMchange}[1]{{\colorbox{purple!70}{#1}}}
%\newcommand{\RMchange}[1]{{\setlength{\fboxsep}{0pt}\colorbox{purple!70}{#1}}}
\newcommand{\RMchange}[1]{#1}

@@

@@

\RMchange{An important aspect of DSLs is to be able to use domain knowledge for their compiler transformations, thus enabling code optimizations that may not be detected by general purpose compilers.
In order to achieve this, P4IRS provides its own IR which can represent elements that are specific to particle computations.
These particle-specific elements in the AST can then be used in P4IRS compilation passes, and at some point during the compilation are lowered, resulting in the low level AST that corresponds to the code implementation that matches the strategies given by the user.}

@@

\RMchange{The P4IRS IR can be constructed in two ways: (a) by mapping of the Python IR generated by triggering the \emph{compile()} method and (b) by explicitly constructing the IR elements in-place.
The former is used for user-defined particle interactions (like the \emph{lennard\_jones} method from \autoref{lst:pairslj}) and the latter is normally used internally by P4IRS to construct the IR for internal routines.}
\autoref{lst:build_neigh} displays an example for constructing the intermediate representation that generates the code that builds the neighbor-lists for each particle in the system.
The \emph{BuildNeighborLists} class is a separate, high-level element from the AST in P4IRS.
During the compilation process, P4IRS performs analyses and transformations to lower this element by using its \emph{lower} method, and decorators such as the \emph{@pairs\_device\_block} are used to determine the target hardware where the code must be executed.

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.75\textwidth]{pairs_architecture.pdf}
  \caption{\RMchange{P4IRS architecture. The user specifies the Particle Computations to be performed, which are compiled to the Python AST and further mapped to the P4IRS IR. Optimization Choices and Simulation Setup methods are used to determine which internal kernels are utilized, and the P4IRS IR is constructed in-place for the Internal Particle Optimization Kernels based on the inputs. Many Compilation Passes are performed to transform the P4IRS IR before triggering the Code Generator, which generates the Optimized Kernels, Data Structures and the PairsSimulation interface class. When whole-program generation is not used, a backend Simulation Code is required, and it uses the PairsSimulation interface to call P4IRS generated kernels. The final Binary is then generated with any backend compiler the user chooses (i.e., GCC, CLANG, ...), and it must be linked with the Runtime Library. The Runtime Library uses the P4IRS Metadata structure to handle different particle data structures and adapt to different simulations.}}
  \label{fig:features}
\end{figure*}

>>>>

% Changes during review
%\usepackage{soul}
%\usepackage{xcolor}
%\DeclareRobustCommand{\RMchange}[1]{{\sethlcolor{cyan}\hl{#1}}}
%\newcommand{\RMchange}[1]{{\colorbox{purple!70}{#1}}}
%\newcommand{\RMchange}[1]{{\setlength{\fboxsep}{0pt}\colorbox{purple!70}{#1}}}
\newcommand{\RMchange}[1]{{\color{blue} #1}} 
%\newcommand{\RMchange}[1]{#1}

@@

\begin{figure*}[htb]
  \centering
  \includegraphics[width=0.75\textwidth]{pairs_architecture.pdf}
  \caption{\RMchange{P4IRS architecture. The user specifies the Particle Computations to be performed, which are compiled to the Python AST and further mapped to the P4IRS IR. Optimization Choices and Simulation Setup methods are used to determine which internal kernels are utilized, and the P4IRS IR is constructed in-place for the Internal Particle Optimization Kernels based on the inputs. Many Compilation Passes are performed to transform the P4IRS IR before triggering the Code Generator, which generates the Optimized Kernels, Data Structures and the PairsSimulation interface class. When whole-program generation is not used, a backend Simulation Code is required, and it uses the PairsSimulation interface to call P4IRS generated kernels. The final Binary is then generated with any backend compiler the user chooses (i.e., GCC, CLANG, ...), and it must be linked with the Runtime Library. The Runtime Library uses the P4IRS Metadata structure to handle different particle data structures and adapt to different simulations.}}
  \label{fig:pairs_arch}
\end{figure*}

@@

\RMchange{This and the following sections focus on describing the architecture from the P4IRS compiler.
\autoref{fig:pairs_arch} depicts a general overview of the P4IRS architecture from the user inputs until the final binary.

An important aspect of DSLs is to be able to use domain knowledge for their compiler transformations, thus enabling code optimizations that may not be detected by general purpose compilers.
In order to achieve this, P4IRS provides its own IR which can represent elements that are specific to particle computations.
These particle-specific elements in the AST can then be used in P4IRS compilation passes, and at some point during the compilation are lowered, resulting in the low level AST that corresponds to the code implementation that matches the strategies given by the user.}

@@

\RMchange{The P4IRS IR can be constructed in two ways: (a) by a mapping from the Python IR generated through the \emph{compile()} method on Python functions and (b) by explicitly constructing the IR elements in-place.
The former is used for user-defined particle interactions (like the \emph{lennard\_jones} method from \autoref{lst:pairslj}) and the latter is normally used internally by P4IRS to construct the IR for internal routines.}
\autoref{lst:build_neigh} displays an example for constructing the intermediate representation that generates the code that builds the neighbor-lists for each particle in the system.
The \emph{BuildNeighborLists} class is a separate, high-level element from the AST in P4IRS.
During the compilation process, P4IRS performs analyses and transformations to lower this element by using its \emph{lower} method, and decorators such as the \emph{@pairs\_device\_block} are used to determine the target hardware where the code must be executed.

@@
